---
title: "ladybird-morphology-body size - hindwing shape"
author: "Heshani Edirisinghe"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load the required packages
library(ggplot2)
library(ade4)
library(cluster)
library(grid)
library(png)
library(colorBlindness)
library(dplyr)

#for wing shape analysis
library(geomorph)
library(shapes)
library(svd)
library(scatterplot3d)
library(rgl)
library(MASS)
library(ape)
library(vegan)

#GLM
library("lme4")
library("lmerTest")
library(MuMIn)
```

#Explore all data for size parameters
``` {r, echo=FALSE, message=FALSE, warning=FALSE}

ladybird_all<- read.csv("data_morpho_1.csv", header=TRUE)

summary(ladybird_all)

```
Notes: There are few missing values in the size variables. For now we will explore with these.

##Exploring distribution of size variables between species

``` {r, echo=FALSE, message=FALSE, warning=FALSE}

# Function to draw faceted graphs for selected continuous variables
draw_faceted_graphs <- function(data, category_var, continuous_vars) {
  
  # Check if the specified variables exist in the dataframe
  if (!(category_var %in% names(data))) {
    stop("Category variable not found in the dataframe.")
  }
  
  for (var in continuous_vars) {
    if (!(var %in% names(data))) {
      stop(paste("Continuous variable '", var, "' not found in the dataframe.", sep = ""))
    }
  }
  
  # Create faceted graphs
  plots <- list()
  for (var in continuous_vars) {
    plots[[var]] <- ggplot(data, aes_string(x = category_var, y = var)) +
      geom_boxplot() 
      labs(title = paste("Faceted by", category_var), x = category_var, y = var)
  }
  
  # Return the list of faceted graphs
  return(plots)
}

#plots of continous variables (size variables) by species
faceted_plots_species <- draw_faceted_graphs(ladybird_all, "species", c("pronotum_width", "two_elytra_points", "width_total", "elytra_length", "head_width"))

# Print the plots
for (plot in faceted_plots_species) {
  print(plot)
}

#plots of continous variables (size variables) by thier origin
faceted_plots_origin <- draw_faceted_graphs(ladybird_all, "origin", c("pronotum_width", "two_elytra_points", "width_total", "elytra_length", "head_width"))

# Print the plots
for (plot in faceted_plots_origin) {
  print(plot)
}


```

Notes: size data are skewed. In total width and distance between 2 elytra points the native species showed a negative skew while introduced species showed a positive skew. Outliers identified - the records were check individually and decided to keep the records as they show the variance in size of different species.

#Principal component analysis for size variables
Since size variables are correlated to each other creating a principal componant analysis. Using nipals algorithm as it allows for missing values. 

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Extracting the continuous variables only
ladybird <- read.csv("data_morpho_1_for_pca.csv")

## 1st NIPALS PCA, extracting 2 components

PCA1 <- nipals (ladybird, nf=2, rec=TRUE, niter=200, tol=1e-09)

paste("summary of 1st NIPALS PCA, extracting 2 components")
PCA1
summary(PCA1)


## 2nd NIPALS PCA, extracting 3 components


PCA2 <- nipals (ladybird, nf=3, rec=TRUE, niter=400, tol=1e-09)
paste("summary of 2nd NIPALS PCA, extracting 3 components")
PCA2
summary(PCA2)



```
Notes: Going to concentrate on the first two principal components (PCA1)

Calculating the eigenvalues of the principal components:

```{r, echo=FALSE, message=FALSE, warning=FALSE}

PCA1$eig

```

Calculating the percentage variance explained by each principal component:
``` {r, echo=FALSE, message=FALSE, warning=FALSE}

(kip <- 100*PCA1$eig/sum(PCA1$eig))

```
Calculating the component loading scores:
``` {r, echo=FALSE, message=FALSE, warning=FALSE}

PCA1$c1

```
Extracting the principal component scores into csv

``` {r, echo=FALSE, message=FALSE, warning=FALSE}

morpho_pca1 <- PCA1$li
#write.csv(PCA1$li, file = "morpho_size_pcscores.csv")

#Merging dataframes
phenovar1 <- data.frame(ladybird_all,morpho_pca1)

#write.csv(phenovar1, "morpho_data_with_pca_no_wingpc.csv")

```

##Graphing the pc scores
``` {r, echo=FALSE, message=FALSE, warning=FALSE}

class(phenovar1$origin)
phenovar1$origin <- as.factor(phenovar1$origin)
group <- phenovar1$species*phenovar1$origin
ggplot(phenovar1, aes(x=V1, y=V2, group = group)) +
  geom_point(aes(shape = origin, color=species), size=8) + 
  #scale_shape_manual(name = "Trend", values = c(1, 2, 1, 2,1,2,1,2,1,2)) +
  #scale_fill_manual(name = "Trend", values = c("blue", "red", "orange", "green", "black", "yellow", "pink", "brown", "purple","white")) +
    #facet_wrap( ~ origin)+
  xlab("Principal component 1") +
  ylab("Principal component 2") +
  theme_classic()+
  #guides(colour = guide_legend(override.aes = list(size=10)), shape = guide_legend(override.aes = list(size=8)) )+
  theme(axis.title.x = element_text(face="bold", size=18),
        axis.text.x = element_text(size=18),
        axis.title.y = element_text(face="bold", size=18),
        axis.text.y = element_text(size=18),
        legend.title=element_blank(),
        legend.text = element_text(size = 18),
        legend.key.size = unit(1,"lines") ,
        strip.text.x = element_text(face="italic", size = 18),
        axis.line.x = element_line(color="black", size = 0.5),
        axis.line.y = element_line(color="black", size = 0.5))
ggsave("plot_introVsNative_species.png", width = 30, height = 15, units = "cm")

```
#Adjusting for colourblindness - for publication
```{r}
path <- system.file("extdata", package = "colorBlindness")
img <- readPNG(file.path("plot_introVsNative_species.png"))
g <- rasterGrob(img, interpolate=TRUE)
grid.draw(cvdPlot(g, 
                  layout = c("origin", "deuteranope", "protanope",
                             "enhanced", "enhanced.deuteranope", 
                             "enhanced.protanope")))

newImg <- replacePlotColor(g)
outfile <- tempfile(fileext = ".png")
png(outfile)
grid.draw(newImg)
dev.off()

outfile

```
#Analysing wing shape data

``` {r, echo=FALSE, message=FALSE, warning=FALSE}

lands <- readland.tps(file = "combined_no_missing_2_final.txt", specID = "ID")
#generate shape variables
gpa.lands <- gpagen(lands)
pca.lands <- plotTangentSpace(gpa.lands$coords, label = TRUE )


wing_pc<- pca.lands$pc.scores
#write.csv(pca.lands$pc.scores,"wing_pcs_final_May_2019.csv")

pca.lands$pc.summary
pca.lands$rotation

pca.lands$pc.summary
pca.lands$pc.scores
pca.lands$pc.shapes
pca.lands$sdev

#plotting wing shape with reference frames for pc1 and pc2

species_names_order <- read.csv("groups.csv")

wing_pc_list <- data.frame(species_names_order, wing_pc)

gp <- as.factor(wing_pc_list$species)

col.gp <- rainbow(length(levels(gp))) # generates a set of different colour over the rainbow spectrum

names(col.gp) <- levels(gp)

col.gp <- col.gp[match(gp, names(col.gp))]

xlab <- paste("Principal Component 1 ", "(", round(pca.lands$pc.summary$importance[2,1]*100, 1), "%)", sep="")
ylab <- paste("Principal Component 2 ", "(", round(pca.lands$pc.summary$importance[2,2]*100, 1), "%)", sep="")

mat <- matrix(c(4,5,0,1,1,2,1,1,3), 3)
layout(mat, widths=c(1,1,1), heights=c(1,1,0.6))# set the size of the rows and columns

# Item 1 to plot, the graph of PC1 vs PC2
par(mar=c(4, 4, 1, 1)) # sets the margins

plot(wing_pc_list$PC1, wing_pc_list$PC2, pch=21, cex=2, bg=col.gp, xlab=xlab, ylab=ylab, asp=T)
legend(-0.075, 0.08, legend= unique(gp), pch=19,  col=unique(col.gp))
ref <- mshape(gpa.lands$coords)

par(mar = c(0,0,0,0)) # sets the margins
plotRefToTarget(ref,pca.lands$pc.shapes$PC1min)

# Item 3
plotRefToTarget(ref,pca.lands$pc.shapes$PC1max)
# Item 4
plotRefToTarget(ref,pca.lands$pc.shapes$PC2max)
# Item 5
plotRefToTarget(ref,pca.lands$pc.shapes$PC2min)

##end here

barplot(pca.lands$pc.summary$sdev^2/sum(pca.lands$pc.summary$sdev^2))
plot(pca.lands$pc.scores[,1:2],pch=15,xlab="PC1",ylab="PC2")
text(pca.lands$pc.scores[,1:2],rownames(pca.lands$pc.scores[,1:2]),pos=4,cex=.5)

library(ggplot2)
ggplot(wing_pc_list, aes(x=PC1, y=PC2, colour= species, label = id )) +
  geom_point() +
  geom_text(aes(label=id), hjust=0, vjust=0 )+
  xlab("PC1") +
  ylab("PC2") +
  theme(axis.title.x = element_text(face="bold", size=14),
        axis.text.x = element_text(size=14),
        axis.title.y = element_text(face="bold", size=14),
        axis.text.y = element_text(size=14),
        legend.title=element_blank(),
        legend.text = element_text(face = "italic"),
        strip.text.x = element_text(face="italic", size = 14),
        axis.line.x = element_line(color="black", size = 0.5),
        axis.line.y = element_line(color="black", size = 0.5))


ggplot(wing_pc_list, aes(x=PC1, y=PC2, colour= species, label = id )) +
  geom_point() +
  xlab("PC1") +
  ylab("PC2") +
  theme(axis.title.x = element_text(face="bold", size=14),
        axis.text.x = element_text(size=14),
        axis.title.y = element_text(face="bold", size=14),
        axis.text.y = element_text(size=14),
        legend.title=element_blank(),
        legend.text = element_text(face = "italic"),
        strip.text.x = element_text(face="italic", size = 14),
        axis.line.x = element_line(color="black", size = 0.5),
        axis.line.y = element_line(color="black", size = 0.5))

```
#GLMM with all size and wing shape PCs
``` {r, echo=FALSE, message=FALSE, warning=FALSE}
all_data <- read.csv("final_morpho_data_2020.csv")

qqp(all_data$size_pc1.t, "norm")

qqp(all_data$size_pc1.t, "lnorm")

body_size1 <- lmer(size_pc1 ~ origin + (1|species) + (1|site), data = all_data)
summary(body_size1)
body_size2 <- lmer(size_pc1 ~ species + (1|site), data = all_data)
summary(body_size2)


body_size1_ <- lmer(elytra_length ~ origin + (1|species) + (1|site), data = ladybird_all)
summary(body_size1_)
body_size2_ <- lmer(elytra_length ~ species + (1|site), data = ladybird_all)
summary(body_size2_)
AIC(body_size2_)
AIC(body_size2)
body_size3 <- lmer(size_pc2 ~ origin + (1|species) + (1|site), data = all_data)
summary(body_size3)
body_size4 <- lmer(size_pc2 ~ species + (1|site), data = all_data)
summary(body_size4)

body_size4_ <- lmer(width_total ~ species + (1|site), data = ladybird_all)
summary(body_size4_)
AIC(body_size4_)
AIC(body_size4)

wing_size1 <- lmer(wing_pc1 ~ species +  (1|site), data = all_data)
summary(wing_size1)
wing_size2 <- lmer(wing_pc2 ~ species +  (1|site), data = all_data)
summary(wing_size2)


paste("AIC of size_pc1 ~ origin model")
AIC(body_size1)
paste("AIC of size_pc1 ~ species model")
AIC(body_size2)
paste("AIC of size_pc2 ~ origin model")
AIC(body_size3)
paste("AIC of size_pc2 ~ species model")
AIC(body_size4)
paste("AIC of wing_pc1 ~ species model")
AIC(wing_size1)
paste("AIC of wing_pc2 ~ species model")
AIC(wing_size2)
```
##Check model fit
``` {r, echo=FALSE, message=FALSE, warning=FALSE}

model_name <- NULL

model_list <- list("body_size1",	"body_size2",	"body_size3","body_size4", "wing_size1", "wing_size2")

model_fit <- data.frame(matrix(ncol=3, nrow = 4))
colnames(model_fit) <- c("model","R2m","R2c")

p=1
for (i in 1:6) {
  
model_name_reset <- model_name
model_name = paste(model_list[p])

    j=1
  model_fit[p,j] <- paste(model_list[p])
  model_fit[p,j+1] <- r.squaredGLMM(get(model_name))[1][1]
  model_fit[p,j+2] <- r.squaredGLMM(get(model_name))[2][1]
model_name <-model_name_reset
  j=1
  p=p+1
  
}

model_fit

model_AIC <- data.frame(matrix(ncol=3, nrow = 6))
colnames(model_AIC) <- c("model","AIC","BIC")

p=1
for (i in 1:6) {
  
model_name_reset <- model_name
model_name = paste(model_list[p])

    j=1
  model_AIC[p,j] <- paste(model_list[p])
  model_AIC[p,j+1] <- AIC(get(model_name))
  model_AIC[p,j+2] <- BIC(get(model_name))
model_name <-model_name_reset
  j=1
  p=p+1
  

}

 
model_AIC

```

##Check linearity of residuals by making a residual plot
```{r echo=FALSE, message=FALSE, warning=FALSE}

for (i in 1:6) {
  
  model_name = paste(model_list[[i]])

plot(resid(get(model_name)) ~ fitted (get(model_name)), main= paste("residual plot",model_name),xlab = paste("fitted",model_name), ylab = paste("resid", model_name))
abline (h=0)
  
}

for (i in 1:6) {
  
model_name = paste(model_list[[i]])
hist(residuals(get(model_name)),main = paste("Histogram of residuals",model_name), xlab = model_name)
    model_name <- model_name_reset
  
  }

```